---
title: "DATA607 - Working with XML and JSON in R"
author: "Nicholas Kunze"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if("RCurl" %in% rownames(installed.packages()) == FALSE) {install.packages("RCurl")}
library(RCurl)
if("jsonlite" %in% rownames(installed.packages()) == FALSE) {install.packages("jsonlite")}
library(jsonlite)
if("XML" %in% rownames(installed.packages()) == FALSE) {install.packages("XML")}
library(XML)
```

## Overview

This assignment goes over the parsing of different data storage formats, specifically HTML, JSON, and XML. I will examine building dataframes from these files and comparing their structure and ease of use.

## JSON

```{r get-data-json}
booksJsonFile <- getURL("https://raw.githubusercontent.com/Kadaeux/DATA607XMLJSON/master/data/books.json")
booksJson <- fromJSON(booksJsonFile)

jsondf <- booksJson$books
jsondf
```


## XML

```{r get-data-xml}
booksXmlFile <- getURL("https://raw.githubusercontent.com/Kadaeux/DATA607XMLJSON/master/data/books.xml")
booksXml <- xmlParse(booksXmlFile)

booksXml
```

Ok, so unlike JSON I still need to deconstruct the returned xml 'object' to build my dataframe. I'll do this by getting the root node, transforming the book nodes into a matrix, then transposing the matrix and reshaping to a dataframe.

```{r xml-build-df}
booksMatrix <- t(xmlSApply(xmlRoot(booksXml), function(x) xmlSApply(x, xmlValue)))

xmldf <- as.data.frame(booksMatrix)
xmldf
```

This way, we still have the issue of authors being smashed together... that'll need to be cleaned. 

## HTML

```{r get-data-html}
booksHtmlFile <- getURL("https://raw.githubusercontent.com/Kadaeux/DATA607XMLJSON/master/data/books.html")
booksHtml <- readHTMLTable(booksHtmlFile,header=TRUE)

htmldf <- booksHtml$`NULL`
htmldf
```

It looks like HTML table parsing uses some part of the table as a name, which I can access the same way I would a df column. Since I did not explicitly name it, it is accessible through 'null'. This is essentially the same basic way I access JSON. 

## Conclusion

HTML and JSON parsing result in very similar dataframes, with JSON producing vectors in nested objects; e.g. authors array turning into vector of names. HTML instead produced a comma delimited list. They were also both exceptionally easy to parse.

XML was much more hands on and annoying to parse, even though I would say it was a more human-readable format than HTML. Its dataframe was much... dirtier, with the authors variable being concatenated without any delimination and therefore requires more effort to handle parsing.